Verification of the CVM algorithm with a Functional Probabilistic Invariant

### 2025-06-20 paper/sec/intro.tex
address PR comments

To show the generality of our technique, we introduce a new variant of the CVM algorithm, where the subsampling step in Line~9 of \cref{alg:cvm} selects a random $m$-subset of $\chi$ instead of independently discarding each element with some probability.
This variant has the benefit that it is \emph{total} (never returns $\bot$) because the second check in Line~11 becomes obsolete.
More interestingly, the resulting variant is \emph{unbiased}, i.e., the expected value is exactly the cardinality of the elements in the stream; this is a new property that neither the original CVM algorithm nor classic algorithms for the distinct elements problem possess.

>>>>

To show the generality of our technique, we introduce a new variant of the CVM algorithm, where the subsampling step in Line~9 of \cref{alg:cvm} selects a random $m$-subset of $\chi$ instead of independently discarding each element with some probability.
This variant has the benefit that it is \emph{total} (never returns $\bot$) because the second check in Line~11 becomes obsolete.
More interestingly, the resulting variant is \emph{unbiased}, i.e., the expected value returned by the algorithm is exactly the cardinality of the elements in the stream; this is a new property that neither the original CVM algorithm nor classic algorithms for the distinct elements problem possess.

### 2025-06-20 paper/sec/semra.tex
address PR comments

\begin{remark}
The theoretically optimal randomized algorithm for distinct elements requires $\bigo( \varepsilon^{-2} \ln \delta + b)$ bits, but it requires more advanced algorithmic techniques. It would not be possible to present using such elementary steps as in \cref{alg:cvm} as it involves computations in finite fields and random walks in expander graphs~\cite{blasiok2020, karayel2023}.
\lipicsEnd\end{remark}

>>>>

\begin{remark}
The asymptotically optimal randomized algorithm for distinct elements requires $\bigo( \varepsilon^{-2} \ln \delta + b)$ bits, but it requires more advanced algorithmic techniques. It would not be possible to present using such elementary steps as in \cref{alg:cvm} as it involves computations in finite fields and random walks in expander graphs~\cite{blasiok2020, karayel2023}.
\lipicsEnd\end{remark}
